### 2023-02-27 data-analytics.qmd
starting data crunching pipeline

---
title: "Data analytics"
format: pdf
---

```{r setup, include=FALSE}
library(tidyverse)
```

This script summarises the data analytical steps for the ICNS2023 paper.

```{mermaid}
%%| fig-width: 6
flowchart LR
  OSN[(Opensky \nNetwork)]
  OSN --> A[Data \nExtraction] --> B(Arrival \nTrajectories)
  B --> C{Decision}
  C --> D[Result one]
  C --> E[Result two]
```

## Opensky Data Extraction

* use/refine code developed for 3Di to download OSN-Monday-Dumps

```{r}
source("./R/osn_dumps.R")
source("./R/osn_read_tardump.R")
source("./R/osn_make_nice_names.R")
```

* OSN dumps have been stored for re-usage in a folder above the project hierarchy
* load a single day, e.g. 1. August 2019
* read in the chosen day
* standardise OSN naming conventions

```{r}
#| message: false
pth_to_dumps <- "../__DATA/OSN-dumps"
ptn <- "2019-08-01"

# list downloaded dumps
fns <- list.files(path = pth_to_dumps, pattern = ptn, full.names = TRUE)

# load list of chosen dumps
ds  <- fns |> 
  purrr::map_dfr(.f = read_osn_tar_file) |> 
  make_nice_names_osn()
```

* we want to study arrivals at certain airports, i.e. 200NM
* let's take London Heathrow (EGLL)
  + conversions done by hand & https://www.fcc.gov/media/radio/dms-decimal

EGLL Lat: 512839N Long: 0002741W   51.4775 -0461389
09L

089.67°

3901 x 50 M	
RWY surface: Asphalt, Grooved
PCN 83/F/A/W/T

SWY surface: Asphalt, Grooved
PCN 83/F/A/W/T

512839.00N 0002905.97W THR 78.6 FT
TDZ 81.3 FT	
27R

269.71°

3901 x 50 M	
RWY surface: Asphalt, Grooved
PCN 83/F/A/W/T

SWY surface: Asphalt, Grooved
PCN 83/F/A/W/T

512839.63N 0002559.82W
151.0 FT	
THR 78.1 FT
TDZ 79.3 FT	
09R

089.68°

3658 x 50 M	
RWY surface: Asphalt, Grooved
PCN 83/F/A/W/T

SWY surface: Asphalt, Grooved
PCN 83/F/A/W/T

512753.25N 0002856.33W
151.0 FT	
THR 75.3 FT
TDZ 76.2 FT	
27L

269.72°

3658 x 50 M	
RWY surface: Asphalt, Grooved
PCN 83/F/A/W/T

SWY surface: Asphalt, Grooved
PCN 83/F/A/W/T

512753.82N 0002602.76W
151.0 FT	
THR 76.8 FT
TDZ 77.8 FT

```{r}
# airport-data.com - server down???   source("./R/get_apt_info.R")
## aip_egll <- get_apt_info("EGLL")

apts <- c("EGLL","EDDF","EHAM","KATL","KJFK","KORD","WSSS")

# default back to ourairports.com
# url <- "https://davidmegginson.github.io/ourairports-data/airports.csv"
# aip_apts <- readr::read_csv(url)
# aip_apts <- aip_apts |> filter(ident %in% apts)
```

~REF, ~LAT, ~LON, ~ELEV, ~SRC
  "EGLL", "OSN-center", 51.4706, -0.461941, 83, "https://opensky-network.org/airport-profile?icao=EGLL"

```{r}
aip_apts <- tribble(
  ~ICAO,   ~REF,         ~LAT,    ~LON,     ~ELEV, ~SRC 
  ,"EGLL", "OSN-center", 51.4706, -0.461941, 83, "https://opensky-network.org/airport-profile?icao=EGLL"
  ,"EGLL", "ARP",        51.4775, -0.461389, NA, "AIP"
  ,"EGLL", "THR-09L",    51.4775, -0.484992, 79 ,"AIP - 512839.00N 0002905.97W THR 78.6 FT"
  ,"EGKK", "OSN-center", 51.148102,-0.190277, 202,"https://opensky-network.org/airport-profile?icao=EGKK"
  ,"EDDF", "OSN-center", 50.033333,8.570556, 364,"https://opensky-network.org/airport-profile?icao=EDDF"
  ,"EHAM", "OSN-center", 52.308601,4.763889, -11, "https://opensky-network.org/airport-profile?icao=EHAM"
  ,"KATL", "OSN-center", 33.6366996,-84.4281005,1026,"https://opensky-network.org/airport-profile?icao=KATL"
  ,"KORD", "OSN-center", 41.9785996,-87.9048004,672, "https://opensky-network.org/airport-profile?icao=KORD"
  ,"KJFK", "OSN-center", 40.6398010,-73.7789002, 13, "https://opensky-network.org/airport-profile?icao=KJFK"
)
write_csv(aip_apts, file = "./data/aip.csv")
aip_apts
```

Check for our study airport whether flights arrive (or depart)

* trajectory of LAT/LON within a certain distance around an airport and below a certain altitude