### 2023-02-27 paper.qmd

---
# template taken from https://github.com/gael-close/ieeetran
#----------------------------------------------------------
title: Arrival Management with Open Data
format:
  ieeetran-pdf:
    keep-tex: true

author:
  - name: Rainer Koelle
    orcid: 0000-0000-0000-0000
    email: rainer.koelle@eurocontrol.int
    note: 1  # workaround i/o linked affiliation
  - name: Second Author
    orcid: 0000-0000-0000-0000
    email: second@example.org
    note: 1
  - name: Third Author
    email: third@example.org
    note: 2
#date: 01-Jan-2022.
affiliations:
  - id: 1
    name: Performance Review Unit, EUROCONTROL
    city: Brussels
    country: Belgium
  - id: 2
    name: Some corporation
    city: New York
    country: USA

abstract: |
  This document is a sample illustrating the Quarto `ieeetran` template.
  It includes the key elements of a scientific articles 
  (references, equations, figures, tables, code, cross references).
  The template enables the generation of IEEE-formatted article from a Jupypter notebook.
keywords: [template, demo]
reference-section-title: References
bibliography: bibliography.bib
note: Extra footnote.
# filters:
#    - latex-environment
environments: [mytable, mywidetable]
---

# Header 1

## header 2

some cool text.

Lorem Ipsum is simply dummy text of the printing and typesetting industry. Lorem Ipsum has been the industry's standard dummy text ever since the 1500s, when an unknown printer took a galley of type and scrambled it to make a type specimen book. It has survived not only five centuries, but also the leap into electronic typesetting, remaining essentially unchanged. It was popularised in the 1960s with the release of Letraset sheets containing Lorem Ipsum passages, and more recently with desktop publishing software like Aldus PageMaker including versions of Lorem Ipsum.