### 2023-02-27 paper.tex

\documentclass[letterpaper, 10 pt, journal, twoside]{IEEEtran}
\IEEEoverridecommandlockouts

% Quarto basics minus two problematic packages
\RequirePackage{scrlfile}
\PreventPackageFromLoading{sidenotes}
\PreventPackageFromLoading{marginnotes}

\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}\usepackage{longtable,booktabs,array}
\usepackage{calc} % for calculating minipage widths
% Correct order of tables after \paragraph or \subparagraph
\usepackage{etoolbox}
\makeatletter
\patchcmd\longtable{\par}{\if@noskipsec\mbox{}\fi\par}{}{}
\makeatother
% Allow footnotes in longtable head/foot
\IfFileExists{footnotehyper.sty}{\usepackage{footnotehyper}}{\usepackage{footnote}}
\makesavenoteenv{longtable}
\usepackage{graphicx}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
% Set default figure placement to htbp
\makeatletter
\def\fps@figure{htbp}
\makeatother

\makeatletter
\makeatother
\makeatletter
\makeatother
\makeatletter
\@ifpackageloaded{caption}{}{\usepackage{caption}}
\AtBeginDocument{%
\ifdefined\contentsname
  \renewcommand*\contentsname{Table of contents}
\else
  \newcommand\contentsname{Table of contents}
\fi
\ifdefined\listfigurename
  \renewcommand*\listfigurename{List of Figures}
\else
  \newcommand\listfigurename{List of Figures}
\fi
\ifdefined\listtablename
  \renewcommand*\listtablename{List of Tables}
\else
  \newcommand\listtablename{List of Tables}
\fi
\ifdefined\figurename
  \renewcommand*\figurename{Figure}
\else
  \newcommand\figurename{Figure}
\fi
\ifdefined\tablename
  \renewcommand*\tablename{Table}
\else
  \newcommand\tablename{Table}
\fi
}
\@ifpackageloaded{float}{}{\usepackage{float}}
\floatstyle{ruled}
\@ifundefined{c@chapter}{\newfloat{codelisting}{h}{lop}}{\newfloat{codelisting}{h}{lop}[chapter]}
\floatname{codelisting}{Listing}
\newcommand*\listoflistings{\listof{codelisting}{List of Listings}}
\makeatother
\makeatletter
\@ifpackageloaded{caption}{}{\usepackage{caption}}
\@ifpackageloaded{subcaption}{}{\usepackage{subcaption}}
\makeatother
\makeatletter
\@ifpackageloaded{tcolorbox}{}{\usepackage[many]{tcolorbox}}
\makeatother
\makeatletter
\@ifundefined{shadecolor}{\definecolor{shadecolor}{rgb}{.97, .97, .97}}
\makeatother
\makeatletter
\makeatother

% Common packages
\usepackage[T1]{fontenc}
\usepackage{lipsum}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{physics}
\usepackage{siunitx}
\usepackage{graphicx}
\usepackage[hyphens]{url}
\usepackage{threeparttable}
\usepackage{xcolor}
\usepackage{float}
\usepackage{xcolor}
\usepackage{booktabs}
\usepackage{makecell}
\usepackage{orcidlink}
% monofont
\usepackage[scaled=0.8]{inconsolata}

%styled tabled generated from pandas
\usepackage{colortbl}
\usepackage{multirow}
\usepackage{stfloats} % https://tex.stackexchange.com/a/324358

% Fix table in 2-column format and enable wide table
% https://tex.stackexchange.com/a/224096
\makeatletter
% My tables
\newenvironment{mytable}[1][htbp]{
    \begin{figure}[#1]
    \footnotesize
    \onecolumn
    \begin{minipage}{0.5\textwidth}}
{
    \end{minipage}
    \twocolumn
    \end{figure}}

\newenvironment{mywidetable}[1][tbp]{
    \begin{figure*}[#1]
    \footnotesize
    \onecolumn
    \begin{minipage}{1.0\textwidth}}
{
    \end{minipage}
    \twocolumn
    \end{figure*}}

\renewenvironment{table*}{\begin{mywidetable}}{\end{mywidetable}\ignorespacesafterend}
\renewenvironment{table}{\begin{mytable}}{\end{mytable}\ignorespacesafterend}    
\makeatother

% Muted text (for filler text)
\newcommand\muted[1]{%
\bgroup
\hskip0pt\color{black!40!}%
#1%
\egroup
}

% color links
\usepackage{hyperref}
\hypersetup{ colorlinks, citecolor=teal, linkcolor=teal, urlcolor=teal}

\begin{document}

% author blocks
\author{
        Rainer Koelle\(^1\)\orcidlink{0000-0000-0000-0000},     Second
Author\(^1\)\orcidlink{0000-0000-0000-0000},     Third Author\(^2\)
    \thanks{\(^1\)Performance Review Unit, EUROCONTROL. \(^2\)Some
corporation.}
    \thanks{Released on: }
    \thanks{Extra footnote..}
}

% title
\title{Arrival Management with Open Data}
\maketitle

% abstract
\begin{abstract}
    This document is a sample illustrating the Quarto \texttt{ieeetran}
    template. It includes the key elements of a scientific articles
    (references, equations, figures, tables, code, cross references).
    The template enables the generation of IEEE-formatted article from a
    Jupypter notebook.
\end{abstract}

% body
\hypertarget{header-1}{%
\section{Header 1}\label{header-1}}

\hypertarget{header-2}{%
\subsection{header 2}\label{header-2}}

some cool text.

Lorem Ipsum is simply dummy text of the printing and typesetting
industry. Lorem Ipsum has been the industry's standard dummy text ever
since the 1500s, when an unknown printer took a galley of type and
scrambled it to make a type specimen book. It has survived not only five
centuries, but also the leap into electronic typesetting, remaining
essentially unchanged. It was popularised in the 1960s with the release
of Letraset sheets containing Lorem Ipsum passages, and more recently
with desktop publishing software like Aldus PageMaker including versions
of Lorem Ipsum.

% done
\end{document}