### 2023-02-27 x_osm-extract-rwys.qmd

---
title: "Extracting OSM data"
format: html
---

```{r setup}
# setup
library(tidyverse)
library(osmdata)
library(sf)
```

```{r}
# OSM query
bb_lonlat <- c(-0.223915,51.141733,-0.155851,51.165738)
apt_icao  <- "EGKK"
apt_name  <- "London Gatwick"

q <- opq(bbox = bb_lonlat) %>% 
  add_osm_feature(
    key = "aeroway"
    ,value =c("aerodrome", "apron", "control_tower", "gate", "hangar"
              ,"helipad", "runway", "taxiway", "terminal") ) %>% 
  osmdata_sf() 
# %>%
#   unique_osmdata()
```

```{r}
g <- ggplot() +
  geom_sf(data = q$osm_polygons
          ,inherit.aes = FALSE
          ,color = "lightblue"
          #,fill  = "lightblue"
          ) +
  geom_sf(data = q$osm_lines %>% filter(aeroway != "runway")
          , color = "grey"
          ) +
  geom_sf(data = q$osm_lines %>% filter(aeroway == "runway"),
          inherit.aes = FALSE,
          color = "black",
          size = 2 #.4
          ,alpha = .8) +
  theme_void() +
  labs(title = "EGKK")
g
```

Extract runway and endpoints

```{r}
rwys_ls <- q$osm_lines |> filter(aeroway == "runway", is.na(runway))

ggplot() + geom_sf(data = rwys_ls)
```

```{r}
rwys_pts <- st_cast(rwys_ls, "POINT")

ggplot() + 
  geom_sf(data = rwys_ls) +
  geom_sf(data = rwys_pts, color = "red")
```

Go for start and endpoints of linestring

```{r}
thrs_pts1 <- lwgeom::st_startpoint(rwys_ls)
thrs_pts2 <- lwgeom::st_endpoint(rwys_ls)

ggplot() + 
  geom_sf(data = rwys_ls) +
  geom_sf(data = thrs_pts1, size = 3, fill = NA) +
  geom_sf(data = thrs_pts2, size = 3, color = "lightgreen" , fill = NA) +
  geom_sf(data = rwys_pts, color = "red")

```

```{r}
my_pts1 <- lwgeom::st_startpoint(rwys_ls[1,])
my_pts2 <- lwgeom::st_endpoint(rwys_ls[1,])

ggplot() + 
  geom_sf(data = rwys_ls[1,]) +
  geom_sf(data = my_pts1, color = "red") + 
  geom_sf(data = my_pts2, color = "blue")
```

```{r}
extract_rwy_linestrings <- function(.osm_sf){
  rwys_ls <- .osm_sf$osm_lines |> filter(aeroway == "runway", is.na(runway))
  message("number of runways: ", nrow(rwys_ls))
  rwys_ls
}

pt_to_latlon <- function(.pt_sf){
  .pt_sf |> 
    sf::st_coordinates() |> 
    dplyr::as_tibble() |> 
    dplyr::rename(LON = X, LAT = Y) 
}

extract_thresholds <- function(.rwy_ls){
  thr1 <- lwgeom::st_startpoint(.rwy_ls) |> pt_to_latlon()
  thr2 <- lwgeom::st_endpoint(  .rwy_ls) |> pt_to_latlon()

  bind_rows(thr1, thr2)
}
```

```{r}
rwys_ls <- extract_rwy_linestrings(q)
extract_thresholds(rwys_ls)
```