### 2023-02-28 x_osm-extract-rwys.qmd

---
title: "Extracting OSM data"
format: html
---

Development version

Data preparatory step: extract runway information from Openstreet Map (OSM).

```{r setup}
# setup
library(tidyverse)
library(osmdata)
library(sf)
```

```{r}
# OSM query
bb_lonlat <- c(-0.223915,51.141733,-0.155851,51.165738)
apt_icao  <- "EGKK"
apt_name  <- "London Gatwick"

q <- opq(bbox = bb_lonlat) %>% 
  add_osm_feature(
    key = "aeroway"
    ,value =c("aerodrome", "apron", "control_tower", "gate", "hangar"
              ,"helipad", "runway", "taxiway", "terminal") ) %>% 
  osmdata_sf() 
# %>%
#   unique_osmdata()
```

```{r}
g <- ggplot() +
  geom_sf(data = q$osm_polygons
          ,inherit.aes = FALSE
          ,color = "lightblue"
          #,fill  = "lightblue"
          ) +
  geom_sf(data = q$osm_lines %>% filter(aeroway != "runway")
          , color = "grey"
          ) +
  geom_sf(data = q$osm_lines %>% filter(aeroway == "runway"),
          inherit.aes = FALSE,
          color = "black",
          size = 2 #.4
          ,alpha = .8) +
  theme_void() +
  labs(title = "EGKK")
g
```

Extract runway and endpoints

```{r}
rwys_ls <- q$osm_lines |> filter(aeroway == "runway", is.na(runway))

ggplot() + geom_sf(data = rwys_ls)
```

```{r}
rwys_pts <- st_cast(rwys_ls, "POINT")

ggplot() + 
  geom_sf(data = rwys_ls) +
  geom_sf(data = rwys_pts, color = "red")
```

Go for start and endpoints of linestring

```{r}
thrs_pts1 <- lwgeom::st_startpoint(rwys_ls)
thrs_pts2 <- lwgeom::st_endpoint(rwys_ls)

ggplot() + 
  geom_sf(data = rwys_ls) +
  geom_sf(data = thrs_pts1, size = 3, fill = NA) +
  geom_sf(data = thrs_pts2, size = 3, color = "lightgreen" , fill = NA) +
  geom_sf(data = rwys_pts, color = "red")

```

```{r}
my_pts1 <- lwgeom::st_startpoint(rwys_ls[1,])
my_pts2 <- lwgeom::st_endpoint(rwys_ls[1,])

ggplot() + 
  geom_sf(data = rwys_ls[1,]) +
  geom_sf(data = my_pts1, color = "red") + 
  geom_sf(data = my_pts2, color = "blue")
```

```{r}
extract_rwy_linestrings <- function(.osm_sf){

 osm_values <- c("osm_id", "aeroway", "ref", "runway", "geometry")

  this_rwys_ls <- .osm_sf$osm_lines |>
    dplyr::select(any_of(osm_values)) 
  message("missing values: ", setdiff(osm_values, names(this_rwys_ls)))

  # fix EDDF no runway-tag!
  if("runway" %in% setdiff(osm_values, names(this_rwys_ls))){
    this_rwys_ls <- this_rwys_ls |> mutate(runway = NA_character_)
  }

  this_rwys_ls <- this_rwys_ls |> 
    dplyr::filter(aeroway == "runway", is.na(runway))
  message("number of runways: ", nrow(this_rwys_ls))
  this_rwys_ls
}

pt_to_latlon <- function(.pt_sf){
  .pt_sf |> 
    sf::st_coordinates() |> 
    dplyr::as_tibble() |> 
    dplyr::rename(LON = X, LAT = Y) 
}

extract_thresholds <- function(.rwys_ls, .drop_geometry = FALSE){
  rwys <- .rwys_ls |> select(aeroway, ref, geometry)
  thr1 <- lwgeom::st_startpoint(.rwys_ls) |> pt_to_latlon() |> rename(THR1_LAT = LAT, THR1_LON = LON)
  thr2 <- lwgeom::st_endpoint(  .rwys_ls) |> pt_to_latlon() |> rename(THR2_LAT = LAT, THR2_LON = LON)

  bind_cols(rwys, thr1, thr2)  # todo: add drop_geometry
}
```

```{r}
rwys_ls <- extract_rwy_linestrings(q)
extract_thresholds(rwys_ls)
```

get thresholds

* sfnetworks has edge_azimuth() function, coerce bearing into degrees
* calculate (rough) offset between bearing and runway identifier & select

```{r}
rq <-  extract_thresholds(rwys_ls)

rq2 <- rq |> 
  sfnetworks::as_sfnetwork() |> 
  sfnetworks::activate("edges") |> 
  dplyr::mutate(  
      BRG = sfnetworks::edge_azimuth()
    , BRG = units::set_units(BRG, "degrees")
    , BRG = (as.numeric(BRG) + 360) %% 360

    , REF1 = stringr::str_split_i(ref, pattern = "/",1)
    , REF2 = stringr::str_split_i(ref, pattern = "/",2)

    , RWY_1 = if_else( abs( (parse_number(REF1) * 10) - BRG ) < 30, REF1, REF2)
    , RWY_2 = if_else( abs( (parse_number(REF2) * 10) - BRG ) > 30, REF2, REF1)
         )
rq2
```

Package output

```{r}
apt <- "EGKK"

# pull out edges/runways 
rq3 <- rq2 |> sfnetworks::activate("edges") |> st_as_sf() |> sf::st_drop_geometry()

rwy_ends <-
  bind_cols(rq3 |> select(REF = RWY_1, LAT = THR1_LAT, LON = THR1_LON, NAME = ref)) |> 
  bind_rows(rq3 |> select(REF = RWY_2, LAT = THR2_LAT, LON = THR2_LON, NAME = ref)) |> 
  mutate(ICAO = apt) |> select(ICAO, everything())
rwy_ends
```

```{r}
ggplot() + 
  geom_sf(data = rwys_ls) +
  geom_point(data = rwy_ends, aes(x = LON, y = LAT)) +
  geom_text(data = rwy_ends, aes(x = LON, y = LAT, label = REF), vjust = -0.5)
```

Let's go for Heathrow

```{}
# fails from github! ### source("./R/xhelper_read_excel_from_url.R")
url   <- "https://github.com/euctrl-pru/airport-dashboard/blob/master/data/APT_BBOX.xlsx"
# github urls follow the pattern: https://raw.github.com/user/repository/branch/file.name
github_download_url <- "https://raw.github.com/euctrl-pru/airport-dashboard/master/data/APT_BBOX.xlsx"

## bboxs <- read_excel_from_url(url)
download.file(github_download_url, destfile = "./data/airport-bbox.xlsx", mode = "wb")
bboxs <- readxl::read_excel("./data/airport-bbox.xlsx")
```

```{r}
query_osm_bbox <- function(.bb_lonlat){
  q <- opq(bbox = .bb_lonlat) %>% 
    add_osm_feature(
      key = "aeroway"
      ,value =c("aerodrome", "apron", "control_tower", "gate", "hangar"
              ,"helipad", "runway", "taxiway", "terminal") ) %>% 
    osmdata_sf() 
}

plot_airport <- function(.osm_apt, .apt=NULL){
  g <- ggplot() +
  geom_sf(data = .osm_apt$osm_polygons
          ,inherit.aes = FALSE
          ,color = "lightblue"
          #,fill  = "lightblue"
          ) +
  geom_sf(data = .osm_apt$osm_lines %>% filter(aeroway != "runway")
          , color = "grey"
          ) +
  geom_sf(data = .osm_apt$osm_lines %>% filter(aeroway == "runway"),
          inherit.aes = FALSE,
          color = "black",
          size = 2 #.4
          ,alpha = .8) +
  theme_void() +
  labs(title = .apt)
 return(g)
}

coerce_bbox_string_to_numeric_vector <- function(.bbox_string){
  .bbox_string |> 
    str_split_1(pattern = ",") |> 
    as.numeric()
}
```

```{r}
egll_bb_lonlat <- bboxs |> 
  filter(AIRPORT == "EGLL") |> pull(BBOX) |> 
  coerce_bbox_string_to_numeric_vector()

osm_apt <- query_osm_bbox(egll_bb_lonlat)

plot_airport(osm_apt)
```

Check for runway ends

```{r}
rwys_ls <- extract_rwy_linestrings(osm_apt)
extract_thresholds(rwys_ls)
```

```{r}
associate_rwy_brg <- function(.thrs_ls){
  .thrs_ls |> 
  sfnetworks::as_sfnetwork() |> 
  sfnetworks::activate("edges") |> 
  dplyr::mutate(  
      BRG = sfnetworks::edge_azimuth()
    , BRG = units::set_units(BRG, "degrees")
    , BRG = (as.numeric(BRG) + 360) %% 360

    , REF1 = stringr::str_split_i(ref, pattern = "/",1)
    , REF2 = stringr::str_split_i(ref, pattern = "/",2)

    , RWY_1 = if_else( abs( (parse_number(REF1) * 10) - BRG ) < 30, REF1, REF2)
    , RWY_2 = if_else( RWY_1 == REF1, REF2, REF1)
         )
}
# rq2 <- rq |> associate_rwy_brg()
# rq2

pull_out_runway_info <- function(.thrs_brgs, .apt = apt){
# pull out edges/runways 
 tmp <- .thrs_brgs |> 
   sfnetworks::activate("edges") |> 
   sf::st_as_sf() |> 
   sf::st_drop_geometry()

 rwy_ends <-
  bind_cols(tmp |> select(REF = RWY_1, LAT = THR1_LAT, LON = THR1_LON, NAME = ref)) |> 
  bind_rows(tmp |> select(REF = RWY_2, LAT = THR2_LAT, LON = THR2_LON, NAME = ref)) |> 
  mutate(ICAO = .apt) |> select(ICAO, everything())
 rwy_ends
}
```

Executed for

* EGLL
* EHAM
* EGKK
* EDDF
* WSSS
* KATL, KORD, KJFK
* SBGR

```{r}
apt <- "SBGR"

# European airports
bboxs <- readxl::read_excel("./data/airport-bbox.xlsx")
# international airports
bboxs_int <- readxl::read_excel("./data/airport-intnl-bbox.xlsx")
bboxs <- bind_rows(bboxs, bboxs_int)

bb_lonlat <- bboxs |> 
  filter(AIRPORT == apt) |> pull(BBOX) |> 
  coerce_bbox_string_to_numeric_vector()

osm_apt <- query_osm_bbox(bb_lonlat)

rwys_ls <- extract_rwy_linestrings(osm_apt)

if(apt == "EDDF"){
  rwys_ls$ref <- gsub(pattern = "^18$", replacement = "18/36", x = rwys_ls$ref)
}

rwy_ends <-  rwys_ls |> 
  extract_thresholds() |> 
  associate_rwy_brg() |> 
  pull_out_runway_info()

write_csv(rwy_ends, file = paste0("./data/rwys-",apt,".csv"))
rwy_ends
```

Visual check

```{r}
plot_airport(osm_apt) +
  geom_point(data = rwy_ends, aes(x = LON, y = LAT)) +
  geom_text( data = rwy_ends, aes(x = LON, y = LAT, label = REF), vjust = -0.3)
```