### 2023-02-28 paper.qmd
initial stab at Quarto paper

---
# template taken from https://github.com/gael-close/ieeetran
#----------------------------------------------------------
title: Arrival Management with Open Data
format:
  ieeetran-pdf:
    keep-tex: true

author:
  - name: Rainer Koelle
    orcid: 0000-0000-0000-0000
    email: rainer.koelle@eurocontrol.int
    note: 1  # workaround i/o linked affiliation
  - name: Second Author
    orcid: 0000-0000-0000-0000
    email: second@example.org
    note: 1
  - name: Third Author
    email: third@example.org
    note: 2
#date: 01-Jan-2022.
affiliations:
  - id: 1
    name: Performance Review Unit, EUROCONTROL
    city: Brussels
    country: Belgium
  - id: 2
    name: Some corporation
    city: New York
    country: USA

abstract: |
  This document is a sample illustrating the Quarto `ieeetran` template.
  It includes the key elements of a scientific articles 
  (references, equations, figures, tables, code, cross references).
  The template enables the generation of IEEE-formatted article from a Jupypter notebook.
keywords: [template, demo]
reference-section-title: References
bibliography: bibliography.bib
note: Extra footnote.
# filters:
#    - latex-environment
environments: [mytable, mywidetable]
---

# Introduction

The contribution of this paper comprise:

-   conceptualisation of sequencing separation for arrival management and development of an open data and open software based implementation of the approach; and

-   use-case application of the developed approach on a subset of airports within the European <!-- and international context -->

# Trajectory-Based Operations - Arrival Management

## header 2

some cool text.

Lorem Ipsum is simply dummy text of the printing and typesetting industry.
Lorem Ipsum has been the industry's standard dummy text ever since the 1500s, when an unknown printer took a galley of type and scrambled it to make a type specimen book.
It has survived not only five centuries, but also the leap into electronic typesetting, remaining essentially unchanged.
It was popularised in the 1960s with the release of Letraset sheets containing Lorem Ipsum passages, and more recently with desktop publishing software like Aldus PageMaker including versions of Lorem Ipsum.

# Data and Conceptual Approach

## Approach

data preparation

-   trajectory data - Opensky Network, weekly downloads
-   airport information - Openstreet Map

data downloaded & script development, data cleaning

## Open Data for Operational Performance Monnitoring

## Arrival Sequencing

sequence separation

# Case Study - Results

## Data Sampling

At the time of writing no global open flight table exists.
For this study, we validated the sample with reference data available to the Performance Review Unit.
Under the EUROCONTROL Performance Review System, airport operators report movement data on a monthly basis [@apdf_v1_2019].

# Conclusions

<!-- what did this paper show -->

This paper aimed at exploring a data-driven approach to measuring arrival management based on open data.

<!-- strength and weaknesses -->

<!-- application -->

<!-- further research -->

<!-- final appraisal -->

# Reproducibility  {.unnumbered}

This paper has been built with the R/RStudio ecosystem.
The draft manuscript and its supporting data preparatory steps are archived at https://github.com/rainer-rq-koelle/paper-2023-ICNS.

The script to download the weekly global datasets are included.
The cleaned trajectory data is stored at: \<\<tbd\>\>.

# Acknowledgment and Disclaimer {.unnumbered}

The authors thank the comments by XX YY ZZ.

The views expressed are the authors' own and do not represent a policy or position of EUROCONTROL <!-- and co-authors organisations -->.