### 2023-02-28 x_trj-apt-landing-association.qmd
associated landing runway (first stab)

---
title: "Trajectory Landing Runway Assignment"
format: html
---

Development version

Data preparatory step: associated a trajectory to a landing runway, including determining characteristic timestamps (e.g. gate/6NM time, landing time).

```{r setup}
# setup
library(tidyverse)
library(sf)

# set default ggplot theme
ggplot2::theme_set(ggplot2::theme_minimal())
```

The identification of the landing direction is based on a simple geospatial heuristic.    
Conceptually, aircraft are aligned with the runway (centerline) before landing.
An aircraft is assigned to a landing runway based on the closeness of its pre-landing positions to the extended runway centerline.

Load test data set (BAWs @ EGLL)

```{r}
tst_fn  <- "tst-egll-baw.csv.gz"
tst_pth <- here::here("data", tst_fn)
baws <- read_csv(tst_pth, show_col_types = FALSE)
```

Data preparation

```{r}
source(here::here("R","coerce_unixepoch_to_datetime.R"))
source(here::here("R", "cast_sf_utils.R"))

# load EGLL environment
rwys_egll <- read_csv("./data/rwys-EGLL.csv", show_col_types = FALSE)
arp_egll  <- read_csv("./data/aip.csv", show_col_types = FALSE) |>
  filter(ICAO == "EGLL", REF == "OSN-center")

baws   <- baws |> coerce_unixepoch_to_datetime()

baw236 <- baws |> filter(FLTID == "BAW236") |> 
  trrrj::filter_positions_at_range(c(arp_egll$LON, arp_egll$LAT), dm = 1, dM = 205, lon = LON, lat = LAT)

baw236_ls <- baw236 |> cast_latlon_to_ls(.drop_coord = FALSE, .group_var = ICAO24)
```

```{r}
ggplot() +
  geom_sf(data = arp_egll |> cast_latlon_to_pts(), size = 3, color = "blue") +
  geom_sf(data = rwys_egll |> cast_latlon_to_pts(), color = "red") +
  geom_sf(data = baw236_ls) +
  labs(title = "Overall trajectory - arrival path")
```

Zoom in on airport area

```{r}
baw236_landing <- baw236 |> 
  trrrj::filter_positions_at_range(c(arp_egll$LON, arp_egll$LAT), 0, 20, lon = LON, lat = LAT)
baw236_landing_ls <- baw236_landing |> cast_latlon_to_ls(.drop_coord = FALSE, .group_var = ICAO24)

ggplot() +
  geom_sf(data = arp_egll |> cast_latlon_to_pts(), size = 3, color = "blue") +
  geom_sf(data = rwys_egll |> cast_latlon_to_pts(), color = "red") +
  geom_sf(data = baw236_landing_ls) +
  labs(title = "Arrival portion")
```

focus on runway

```{r}
# setup sf objects
arp_pt   <- arp_egll |> cast_latlon_to_pts()
rwys_pts <- rwys_egll |> cast_latlon_to_pts()
rwys_ls  <- rwys_egll |> 
  group_by(NAME) |> 
  group_modify(.f = ~ cast_latlon_to_ls(.) ) |> 
  st_as_sf()

# utility function to create centerline
rwy_ctr_line <- function(.rwy_df = aip, .ctrl_length = 10000){
  df <- .rwy_df %>% 
    filter(REF != "ARP") %>% 
    select(REF, NAME, LAT, LON) %>% 
    group_by(NAME) %>% 
    mutate( LAT2 = case_when(!is.na(lag(LAT))  ~ lag(LAT)
                            ,!is.na(lead(LAT)) ~ lead(LAT))
           ,LON2 = case_when(!is.na(lag(LON))  ~ lag(LON)
                            ,!is.na(lead(LON)) ~ lead(LON))
          ) %>% 
  # calculate "reverse" runway bearing with geosphere::bearingRhumb
  mutate( RBRG  = geosphere::bearingRhumb(p1 = cbind(LON2, LAT2), p2 = cbind(LON, LAT))
         )

  # determine "endpoint" of extended centerline at d = 10000 meters
  tmp <- with(df, geosphere::destPointRhumb(cbind(LON, LAT), b= RBRG, d = .ctrl_length)) %>% 
    as_tibble() %>% rename(LON3 = lon, LAT3 = lat)

  # combine and return
  df <- df %>% bind_cols(tmp)
  return(df)
}

# utility function to coerce extension points
rwy_ctr_line_pts <- function(.rwy_ctr_line, .debug = FALSE){
  # could not get pivot_longer work with multiple cols
  tmp1 <- .rwy_ctr_line %>% select("REF":"LON")

  # include opposite runway threshold
  tmp2 <- .rwy_ctr_line %>% select("REF","NAME", "LAT2","LON2") %>%
    rename(LAT = LAT2, LON = LON2)

  # centerline end point determined beforehand
  tmp3 <- .rwy_ctr_line %>% select("REF","NAME", "LAT3","LON3") %>%
    rename(LAT = LAT3, LON = LON3)

  if(.debug == TRUE){ 
    df <- bind_rows(tmp1, tmp2, tmp3)
  }else{
    df <- bind_rows(tmp1, tmp3)
  }

  df <- df %>% arrange(REF, NAME)
  return(df)  
}

cast_rwy_ctr_line_ls <- function(.rwys_df){
  .rwys_df |> 
  rwy_ctr_line() |> 
  rwy_ctr_line_pts() |> 
  cast_latlon_to_pts() |> 
  cast_pts_to_ls(.group_var = REF)
}
rwy_centerline_ls <- cast_rwy_ctr_line_ls(rwys_egll)
```

```{r}
ggplot() +
  geom_sf(data = arp_pt, size = 3, color = "blue") +
  geom_sf(data = rwys_ls) +
  geom_sf(data = rwys_pts, color = "red") +
  geom_sf(data = rwy_centerline_ls, color = "green") +
  geom_sf_text(data = rwy_centerline_ls, aes(label = REF)) +
  geom_sf(data = baw236_landing_ls)
```

check proximity on final

```{r}
# spatial join (aka "overlay" with "within")
# to identify most likely ~ closest runway (extended centerline-string)
estimate_rwy <- function(.trj_pts, .rwy_ctrl_ls, .ctrl_offset = 1000){
  # stop if not
  tmp <- st_join( .rwy_ctrl_ls, .trj_pts
                 , join = st_is_within_distance, dist = .ctrl_offset) %>%
  st_drop_geometry() %>%        # keep just the "spatially joined" dataframe
  na.omit() %>%                 # eliminate no hits
  group_by(REF) %>%
  summarise(N = n(), .groups = "drop") %>%
  mutate( TOT   = sum(N, na.rm = TRUE)       # na.rm should not be necessary
         ,TRUST = N / TOT)                   # share in case several runways

return(tmp)
}
```

```{r}
baw236_final_approach <- baw236 |> 
  trrrj::filter_positions_at_range(c(arp_egll$LON, arp_egll$LAT), 0, 10, lon = LON, lat = LAT)
# baw236_landing_ls <- baw236_landing |> cast_latlon_to_ls(.drop_coord = FALSE, .group_var = ICAO24)
ggplot() +
  geom_sf(data = arp_pt, size = 3, color = "blue") +
  geom_sf(data = rwys_ls) +
  geom_sf(data = rwys_pts, color = "red") +
  geom_sf(data = rwy_centerline_ls, color = "green") +
  geom_sf_text(data = rwy_centerline_ls, aes(label = REF)) +
  geom_sf(data = baw236_final_approach |> cast_latlon_to_ls())
```

```{r}
estimate_rwy(rwy_centerline_ls, baw236_final_approach |> cast_latlon_to_pts())
```

Try to understand why / how the closeness is calculated.
Current implementation shows N=1 ?!? while ok, it should be multiple hits.