### 2023-03-02 x_trj-classification-segmentation.qmd
trajectory classification

---
title: "Trajectory Classification"
format: html
---

```{r setup}
# setup
library(tidyverse)
library(sf)

source(here::here("R", "level_segments_check.r"))

# set default ggplot theme
ggplot2::theme_set(ggplot2::theme_minimal())
```

DEVELOPMENT VERION

# Overview

We conceptualise

* a **trajectory** as the time-ordered sequence of 4D positions of a flight;
* a **flight** is an air transport movement from a departure aerodrome to its destination aerodrome;
* flights could include segments with specific operations (e.g. military flights: military exercise area, refueling, all flights: flight maneuvers, holding, low approach/touch-and-go)
* outside of these specific operations, flights progress through phases: (i) parking, (ii) taxi-out, including take-off roll, (iii) initial climb-out, (iv) climb-out, (v) en-route, (vi) decent, (vii) final approach, (viii) taxi-in to parking position.

download sample = state vector data (4D positions)

==> split flights into DA(I)O ~ ODA

```{mermaid}
%%| fig-width: 6
flowchart LR
  OSN[(Opensky \nNetwork)] 
  OSN --> A[Data \nExtraction] 
  A --> FP(Flight Phases)
  A --> B(Arrival \nTrajectories)
  FP -.- B
```

# Model Development

Load test data set (BAWs @ EGLL)

```{r}
tst_fn  <- "tst-egll-baw.csv.gz"
tst_pth <- here::here("data", tst_fn)
baws <- read_csv(tst_pth, show_col_types = FALSE)

if(class(baws$TIME) == "numeric") baws <- baws |> coerce_unixepoch_to_datetime()

# clean trajectory
baws <- baws |> Alt_Outlier_filter(.alt_var = ALT_B)
```

TODO = outlier filter works well when airborne, ground jumps not picked up

```{r}
baws |> 
ggplot() +
  geom_point(aes(x = TIME, y = ALT_B, color = FLTID))
```

## Flight Table and Milestone Table

A flight can be characterised by a series of milestones.    
Specific milestones (and supporting data) can be extracted in form of a flight table.

Unique flight identification:

An aircraft (identifiable by its ICAO24 address) can perform multiple flights on a given day.
Flights may enter/leave the coverage area.
Conceptually - and within the studied context of flights operating at European airports / night flying restrictions - the associated of the date of flight and ICAO24 address including the FLTID can serve as a unique key.

Milestone table allows for combining flights on day n and n+1, as characteristic milestones and their time offset allow for an association (and subsequent cleaning of the UID - not part of this study).

### Identify "flights"

On the basic level, flights are detected between START (first seen) and END (last seen) of an ICAO24 - typically - using a single FLTID

```{r}
identify_legs <- function(.trjs, .uid = ICAO24, .max_step = 300){
  .trjs |> 
    dplyr::group_by({{.uid}}) |> 
    dplyr::arrange(TIME) |> 
    dplyr::mutate(
       STEP = difftime(TIME, lag(TIME, default = dplyr::first(TIME))) |> as.numeric()
      ,LEG  = cumsum(STEP > .max_step)
      ,N_ID = length(unique(FLTID))
      ) |> 
    ungroup()
} 
make_uid <- function(.trjs_legs){
  .trjs_legs |> 
    dplyr::mutate(
       DOF = lubridate::date(TIME) 
      ,UID = paste(ICAO24, DOF, LEG, sep = "-")
      ) |> 
    dplyr::select(UID, ICAO24, TIME, LAT, LON, ALT_B, FLTID, everything())
}

init_flight_table <- function(.trjs_uid){
  this_flt <- .trjs_uid  |> 
    dplyr::summarise(
       UID = unique(UID)
      ,FLTID = unique(FLTID) 
      ,N   = n()
      ,FIRST = min(TIME), LAST = max(TIME), DUR = difftime(LAST, FIRST, units = "min")
      ,.by = UID
      ) 
}

init_milestone_table <- function(.trj_uid){
  this_mst <- .trj_uid  |> 
    dplyr::slice(c(1,n())) |> 
    dplyr::mutate(MST = c("START","STOP")) |> 
    dplyr::select(ICAO24, FLTID, TIME, LAT, LON, ALT_B, MST)
}
```

```{r}
mst <- baws |> identify_legs() |> make_uid() |> 
  group_by(UID) |> 
  group_modify(.f = ~ init_milestone_table(.x))
```

DETECT LEVEL SEGMENTS
TOC / TOD

```{r}
baws |> filter(FLTID == "BAW236") |> check_level_flight(.alt_var = ALT_B) |> ggplot() + geom_point(aes(x = TIME, y = ALT_B, color = as.factor(LEVEL)))
```

DETECT AIRPORT BOX

load airport environment

```{r}
# load EGLL environment
rwys_egll <- read_csv("./data/rwys-EGLL.csv", show_col_types = FALSE)
arp_egll  <- read_csv("./data/aip.csv", show_col_types = FALSE) |>
  filter(ICAO == "EGLL", REF == "OSN-center")

rwys_egll_pts <- rwys_egll |> cast_latlon_to_pts()
rwys_egll_pts_buf <- rwys_egll_pts |> sf::st_buffer(dist = 500)
rwys_egll_box <- rwys_egll_pts_buf |> sf::st_union() |> sf::st_convex_hull()
```

```{r}
ggplot() +
  geom_sf(data = rwys_egll_pts) +
  geom_sf(data = rwys_egll_pts_buf, color = "blue", fill = NA) +
  geom_sf(data = rwys_egll_box, color = "lightgreen", fill = NA)
```