### 2023-03-20 x_bra-check.qmd

---
title: "Brazil"
format: html
---

```{r}
library(tidyverse)
```

Unfortunately https://aisweb.decea.mil.br/ was down (20.03.).    
Google search offered an AIP pdf from 2017 (hope that no major changes took place - check with DECEA).

Get airport information from AIP

```{r}
knitr::include_graphics("./figures/aip-snapshot-SBGR.png")
```

use measurements package to calculate decimal degrees (rounded to 6 digits)
measurements::conv_unit("046 28 23", "deg_min_sec", "dec_deg")

SBGR: ARP 232608S / 0462823W

bboxfinder - box lng/lat - min/max
bounding box SBGR: -46.500975,-23.460242,-46.435372,-23.416276

```{r}
rwys_sbgr <- tribble(
   ~ICAO  , ~REF , ~LAT      , ~LON      , ~NAME    , ~ELEV
  , "SBGR", "ARP", -23.435555, -46.473056, "ARP"    , 2461
  , "SBGR", "O9L", -23.434281, -46.483361, "09L/27R", 2444
  , "SBGR", "27R", -23.424894, -46.448619, "09L/27R", 2440
  , "SBGR", "09R", -23.438797, -46.487027, "09R/27L", 2450
  , "SBGR", "27L", -23.431189, -46.458855, "09R/27L", 2445
)

#write_csv(rwys_sbgr, "./data/rwys-SBGR.csv")
```

```{r}
#sbgr_arp   <- c(rwys_sbgr$LON, rwys_sbgr$LAT)
sbgr_200NM <- trrrj::polygon_at_distance(c(8.570556, 50.03333), d = 205)
```

## Load OSN data - Monday dump

```{r}
# load helper scripts to download and store OSN data
source("./R/osn_dumps.R")
source("./R/osn_read_tardump.R")
source("./R/osn_make_nice_names.R")
```
```{r}
#| message: false
pth_to_dumps <- "../__DATA/OSN-MondayDumps"
ptn <- "2022-05-09"

# list downloaded dumps
fns <- list.files(path = pth_to_dumps, pattern = ptn, full.names = TRUE)

# load list of chosen dumps
bra  <- fns |> 
  purrr::map_dfr(.f = read_osn_tar_file) |> 
  make_nice_names_osn()
```