### 2023-03-20 x_data-crunching.qmd

---
title: "Data Preparation"
format: pdf
---

DEVELOPMENT VERSION

```{r}
#| message: false
library(tidyverse)
library(sf)

# set default ggplot theme
ggplot2::theme_set(ggplot2::theme_minimal())
```

## Overview

This Quarto-markdown document lists the data preparatory steps.   
The following is developed for London Heathrow.

1. Load OSN data
2. load target airport(s)

## Load OSN data

```{r}
source("./R/osn_dumps.R")
source("./R/osn_read_tardump.R")
source("./R/osn_make_nice_names.R")
source("./R/coerce_meter_to_feet.R")
```

```{r}
#| message: false
#| warning: false
pth_to_dumps <- "../__DATA/OSN-dumps"
ptn <- "2019-08-01"

# list downloaded dumps
fns <- list.files(path = pth_to_dumps, pattern = ptn, full.names = TRUE)

# load list of chosen dumps
ds  <- fns |> 
  purrr::map_dfr(.f = read_osn_tar_file) |> 
  make_nice_names_osn() |> 
  mutate(ALT_B = coerce_meter_to_feet(ALT_B))
```

## Load target airport

```{r}
source(here::here("R", "cast_sf_utils.R"))
source(here::here("R", "cast_sf_rwy.R"))

# load EGLL environment
rwys_egll <- read_csv("./data/rwys-EGLL.csv", show_col_types = FALSE)
arp_egll  <- read_csv("./data/aip.csv", show_col_types = FALSE) |>
  filter(ICAO == "EGLL", REF == "OSN-center")

# cast sf
arp_pt    <- arp_egll  |> cast_latlon_to_pts(.drop_coord = FALSE) 
rwys_pts  <- rwys_egll |> cast_latlon_to_pts()
rwys_ls   <- rwys_egll |> cast_rwy_ls()

rwy_centerline_ls <- cast_rwy_ctr_line_ls(rwys_egll)
rwy_box           <- airport_threshold_box(rwys_egll)
```

```{r}
apt_arrival_plot <- ggplot() +
  geom_sf(data = rq) +
  geom_sf(data = rwy_box) +
  geom_sf(data = arp_pt, size = 3, color = "blue") +
  geom_sf(data = rwys_ls) +
  geom_sf(data = rwys_pts, color = "red") +
  geom_sf(data = rwy_centerline_ls, color = "green") +
  geom_sf_text(data = rwy_centerline_ls, aes(label = REF)) 
apt_arrival_plot
```

apt_arrival_plot +
  geom_sf(data = baw236_landing_ls)