### 2023-03-25 x_data-crunching.qmd
getting towards completing arrival trajectory extraction

---
title: "Data Preparation"
format: pdf
---

DEVELOPMENT VERSION

```{r}
#| message: false
library(tidyverse)
library(sf)
library(patchwork)

# set default ggplot theme
ggplot2::theme_set(ggplot2::theme_minimal())
```

## Overview

This Quarto-markdown document lists the data preparatory steps.   
The following is developed for London Heathrow.

1. Load OSN data
2. load target airport(s)

## Load OSN data

```{r}
source("./R/osn_dumps.R")
source("./R/osn_read_tardump.R")
source("./R/osn_make_nice_names.R")
source("./R/coerce_meter_to_feet.R")
source("./R/altitude_outlier_filter.R")
source("./R/trj_fix_altitude.R")
source("./R/spatial_subset_jormungand.R")
```

```{r}
#| message: false
#| warning: false
pth_to_dumps <- "../__DATA/OSN-dumps"
ptn <- "2019-08-01"

# list downloaded dumps
fns <- list.files(path = pth_to_dumps, pattern = ptn, full.names = TRUE)

# load list of chosen dumps
ds  <- fns |> 
  purrr::map_dfr(.f = read_osn_tar_file) |> 
  make_nice_names_osn() |> 
  mutate(across(.cols = c("ALT_B","ALT_G"), .fns = coerce_meter_to_feet))
```

## some trajectory cleaning

we study airborn part, i.e. from take-off to landing

* departing flights: velocity > 50 m/s
* landing flights cannot have altitude above 5000ft, velocity < 90

```{r}
ds  |> 
  filter(ICAO24 == "3c6498") |> 
  filter(velocity > 50) |> 
  Alt_Outlier_filter(.alt_var = ALT_B) |> 
  ggplot() + 
    geom_point(aes(x = TIME, y = ALT_B, color = velocity < 90 & ALT_B > 5000)) + 
    geom_point(aes(x = TIME, y = ALT_G), color = "green") +
  theme(legend.position = "top")
```

```{r}
ds  |> 
  filter(ICAO24 == "3c6498") |> 
  fix_altitude_heuristic() |> 
  Alt_Outlier_filter(.alt_var = ALT_B) |> 
  ggplot() + 
    geom_point(aes(x = TIME, y = ALT_B, color = velocity < 90 & ALT_B > 5000)) + 
    geom_point(aes(x = TIME, y = ALT_G), color = "green") +
  theme(legend.position = "top")
```

## Load target airport

```{r}
source(here::here("R", "cast_sf_utils.R"))
source(here::here("R", "cast_sf_rwy.R"))

# load EGLL environment
rwys_egll <- read_csv("./data/rwys-EGLL.csv", show_col_types = FALSE)
arp_egll  <- read_csv("./data/aip.csv", show_col_types = FALSE) |>
  filter(ICAO == "EGLL", REF == "OSN-center")

# cast sf
arp_pt    <- arp_egll  |> cast_latlon_to_pts(.drop_coord = FALSE) 
rwys_pts  <- rwys_egll |> cast_latlon_to_pts()
rwys_ls   <- rwys_egll |> cast_rwy_ls()

rwy_centerline_ls <- cast_rwy_ctr_line_ls(rwys_egll)
rwy_box           <- airport_threshold_box(rwys_egll)
airport_box       <- airport_centerline_box(rwys_egll)

# for later
jormungand <- trrrj::polygon_at_distance(c(arp_egll$LON, arp_egll$LAT), d = 205) |> 
  sf::st_as_sf() |> rename(geometry = x) |> mutate(NAME = "EGLL-205")
```

Check some of the functions and shapes

```{r}
apt_arrival_plot <- ggplot() +
  geom_sf(data = airport_box) +
  geom_sf(data = arp_pt, size = 3, color = "blue") +
  geom_sf(data = rwys_ls) +
  geom_sf(data = rwys_pts, color = "red") +
  geom_sf(data = rwy_centerline_ls, color = "green") +
  geom_sf_text(data = rwy_centerline_ls, aes(label = REF)) 
apt_arrival_plot
```

```{r}
icao24_in_polygon <- function(.ds, .poly, .id = ICAO24){
  # determine bounding box for polygon
  this_box <- sf::st_bbox(.poly)
  # pre-filter with dplyr - cut bounding box
  icao24s <- .ds |> 
    dplyr::filter(
       LAT > this_box$ymin, LAT < this_box$ymax
      ,LON > this_box$xmin, LON < this_box$xmax, 
    )
  # get ids seen
  icao24s <- icao24s |> 
    dplyr::summarise(
      N = n(), N_ID = length(unique(FLTID))
      , MIN_TIME = min(TIME,na.rm = TRUE)
      , MIN_TIME_ALT = ALT_B[which(TIME == MIN_TIME)]
      , MAX_TIME = max(TIME, na.rm = TRUE)
      , MAX_TIME_ALT = ALT_B[which(TIME == MAX_TIME)]
      , MIN_ALT = min(ALT_B, na.rm = TRUE)
      , MED_ALT = median(ALT_B, na.rm = TRUE)
      , MAX_ALT = max(ALT_B, na.rm = TRUE)
      , .by = {{.id}})
}
```

```{r}
seen_at_airport <- icao24_in_polygon(ds, airport_box)
seen_at_airport
```

```{r}
source("~/RProjects/paper-2023-ICNS/R/trj_identify_legs.R")
source("~/RProjects/paper-2023-ICNS/R/coerce_unixepoch_to_datetime.R")
source("~/RProjects/paper-2023-ICNS/R/trj_fix_fltid.R")
source("~/RProjects/paper-2023-ICNS/R/add_distance_rowwise.R")
source("~/RProjects/paper-2023-ICNS/R/trj_fix_altitude.R")
source("~/RProjects/paper-2023-ICNS/R/heuristic_DAO.R")

# helper function to check missing FLTID or gaps
check_leg_fltid <- function(.trjs_with_legs, ...){
  .trjs_with_legs |> 
    dplyr::summarise(
        N = n(), N_ID = length(unique(FLTID))
      , IDs = paste(unique(FLTID), collapse = "-")
      , NAs = sum(is.na(FLTID))
      , .by = c("ICAO24","LEG")
      )
}
```

```{r}
rq <- ds |> 
  # subset for ICAO24s seen at the study airport
  filter(ICAO24 %in% seen_at_airport$ICAO24) |> 
  spatial_subset_jormungand(jormungand) |> 
  # human readable TIME
  coerce_unixepoch_to_datetime() |>
  # clean ALT and trim for airborne part with heuristic fix
  Alt_Outlier_filter(.alt_var = ALT_B) |> 
  fix_altitude_heuristic() |> 
  # assign leg for each airborne segment
  identify_trajectory_legs() |> 
  # force single FLTID here to construct niche UIDs
  group_by(ICAO24, LEG) |> fix_invalid_fltid() |> ungroup() |> 
  # construct UID
  mutate(UID = paste(ICAO24, LEG, sep = "-"))

# check one more time for airport
seen_at_airport2 <- icao24_in_polygon(rq |> sf::st_drop_geometry(), airport_box, UID) |> 
  heuristic_DAO()
seen_at_airport2
```

```{r}
viz_check_DAO <- function(.trj, .apt_box = airport_box){
  plt <- .trj |> 
    ggplot() + 
      geom_sf(data = .apt_box) + 
      geom_point(data = .trj, aes(x = LON, y = LAT, color = TIME)) + 
      coord_sf(xlim = c(-1, 0), ylim = c(51.3, 51.6), expand = FALSE)
  return(plt)
}

viz_check_vertical <- function(.trj){
  plt <- .trj |> 
    ggplot() +
    geom_point(aes(x = TIME, y = ALT_B))
  return(plt)
}

picks <- seen_at_airport2$UID[1:9]

rq |> filter(UID %in% picks) |> viz_check_DAO() + facet_wrap(.~UID)
```

per arrival trajectory (iterate, group modify)

* subset to 200NM (done - jormungand)
```{r}
fig <- ggplot() + geom_sf(data = arr_trjs |> cast_pts_to_ls(.group_var = UID)) +
  labs(subtitle = paste("EGLL arrivals", ptn, sep = "-")) +
  theme(plot.title.position = "plot")
#ggsave(plot = fig, filename = "./figures/EGLL-arrivals-single-day.png", dpi = 300, height = 5, width = 5)
fig
```

* initialise mst (done - initialise milestone table function on filtered trajectories)
* assign landing runway & landing time
* calculate characteristics
  + time-to-go
  + distance-to-go
  + crossings (e.g for ASMA)

* save out mst/enriched trajectory

```{r}
source("~/RProjects/paper-2023-ICNS/R/milestone_functions.R")
arr_trjs <- rq |> filter(UID %in% (seen_at_airport2 |> filter(DAO == "A") |> pull(UID)) )
mst <- initialise_milestone_table(arr_trjs)
mst
```

landing runway assignment

```{r}
source("~/RProjects/paper-2023-ICNS/R/trj_landing_runway.R")

# test for one
arr_rwy <- arr_trjs |> filter(UID == "00b1f9-0") |> assign_landing_runway()
arr_rwy

arr_rwys <- arr_trjs |> assign_landing_runway()
arr_rwys
```

```{r}
arr_rwys |> ggplot() + geom_histogram(aes(x = TRUST), bindwidth = 2)
```

# check for some fringe cases

```{r}
fringe_picks <-  arr_rwys |> filter(between(TRUST, 0.4,0.7))

viz_check_DAO(arr_trjs |> filter(UID %in% fringe_picks$UID)) + facet_wrap(.~ UID)
```

Potential to remove = no landing with TRUST < 0.7

* assign landing runway & landing time

do for one/some

> arr_rwys
# A tibble: 733 × 6
# Groups:   UID [733]
   UID      REF       N   TOT TRUST COMMENT                                                      
   <chr>    <chr> <int> <int> <dbl> <chr>                                                        
 1 00b1f9-0 27L      13    16 0.812 landing runway: 27L trust-Level (hits: 13): 0.8125           
 2 01013c-0 27L      11    14 0.786 landing runway: 27L trust-Level (hits: 11): 0.785714285714286
 3 0101bb-0 27R      12    15 0.8   landing runway: 27R trust-Level (hits: 12): 0.8              
 4 0200cb-0 27R      12    15 0.8   landing runway: 27R trust-Level (hits: 12): 0.8              
 5 04014c-0 27L      12    15 0.8   landing runway: 27L trust-Level (hits: 12): 0.8              
 6 04c15b-0 27R      11    14 0.786 landing runway: 27R trust-Level (hits: 11): 0.785714285714286
 7 06a0e6-0 27R      12    15 0.8   landing runway: 27R trust-Level (hits: 12): 0.8              
 8 06a11f-0 27R      12    15 0.8   landing runway: 27R trust-Level (hits: 12): 0.8              
 9 06a144-0 27L      13    16 0.812 landing runway: 27L trust-Level (hits: 13): 0.8125           
10 06a149-0 27L      13    16 0.812 landing runway: 27L trust-Level (hits: 13): 0.8125

```{r}
arr_trjs |> filter(UID %in% c("00b1f9-0","01013c-0","0101bb-0","0200cb-0")) |> 
  viz_check_DAO() + facet_wrap(.~UID)
```

```{r}
arr_trjs |> filter(UID %in% c("00b1f9-0")) |> 
  ggplot(
    aes(  x = TIME
        , y = ALT_B
        , color = between(ALT_B, arp_egll$ELEV + 100, arp_egll$ELEV + 1000)
        )
    ) + 
  geom_point() + labs(color = "Approach")
```

```{r}
source("~/RProjects/paper-2023-ICNS/R/cast_sf_network.R")

arp  <- arp_egll
rwys <- rwys_egll

final <- arr_trjs |> filter(UID %in% c("00b1f9-0")) |>
  filter(ALT_B < arp$ELEV + 2000)

viz_check_vertical(final)
```

```{r}
source("~/RProjects/paper-2023-ICNS/R/identify_threshold_overhead.R")
source("~/RProjects/paper-2023-ICNS/R/add_bearing_to_target.R")

ldg_rwy <- final |> assign_landing_runway(.arp = arp) |> 
  inner_join(rwys, by = "REF")
my_thr  <- ldg_rwy |> select(UID, REF, LAT, LON, COMMENT) |> 
  cast_latlon_to_pts(.drop_coord = FALSE)

# check if we have a threshold overflight
ovr_thr <- identify_threshold_overhead(final, my_thr)
if(nrow(ovr_thr) > 0){
  message("THRESHOLD")
  eldt  <- ovr_thr |> filter(DIST_THR == min(DIST_THR))
  final <- final |> mutate(MST_LABEL = "ALDT") 
}else{
  message("NEED TO WORK ON LANDING TIME")
}

viz_check_vertical(final) + viz_check_DAO(final)
```

Lets' work this out.

```{r}
source("~/RProjects/paper-2023-ICNS/R/assign_threshold_time.R")

# we have identified an arrival trajectory
trj_arr <- arr_trjs |> filter(UID %in% "0200cb-0") |> 
#trj_arr <- arr_trjs |> filter(UID %in% c("00b1f9-0")) |> 
  assign_threshold_time(my_thr)
```

look at function ahead ... started adding threshold --> complement/fill columns & calc landing time

```{r}
viz_check_vertical(trj_arr)
```

PICK SOME ARRIVALS FOR GETTING CODE IN PLACE

```{r}
# derived from validation data set, binned per 15 mins per arrival runway
# 27L 06:30 - 0645
peak_wave <- c("BAW377", "EWG6VP", "BAW11J", "ETD89G", "AIC161", "AFR82FE", "BAW236","DLH3WU","VIR138M","EWG30N")

peak_trjs <- arr_trjs |> filter(FLTID %in% peak_wave)

viz_check_DAO(peak_trjs) + viz_check_vertical(peak_trjs |> filter(TIME > lubridate::ymd_hm("2019-08-01 06:20")))
```

```{r}
# run loop with group_modify()
# create a new UID for the looping as group_modify cannot accept UID in payload
# TODO - write functions independent of UID or handle UID outside calculation.
study_trjs <- peak_trjs |> 
  mutate(LUID = UID) |> group_by(LUID) |> 
  group_modify(.f = ~assign_threshold_time(.x,.my_thr = my_thr)) |> 
  ungroup() |> select(-LUID)
```

```{r}
ggplot(data = study_trjs) +
  geom_point(aes(x = TIME, y = ALT_B)) +
  geom_vline(xintercept = study_trjs$TIME[which(study_trjs$MST == "ALDT")])
```

first pair

```{r}
# "BAW377","EWG6VP","BAW11J"
inter_arrival <- study_trjs |> filter(MST == "ALDT") |> group_by(MST_REF) |> arrange(TIME, .by_group = TRUE) |> 
  mutate(
     FOLLOWER = lead(UID, default = NA_character_)
    ,SEP_TIME = difftime(lead(TIME, default = NA), TIME, unit = "sec")
    )
inter_arrival
```

```{r}
p1 <- inter_arrival |> ggplot() +
  geom_step(aes(x = TIME, y = SEP_TIME))
p2 <- inter_arrival |> ggplot() +
  geom_boxplot(aes(x = MST_REF, y = SEP_TIME))
p1 + p2
```

OK - FUCKED UP HERE 
NEED to catch improper trajectories that throw error ==> improve cleaning
wrap look in safely/possibly ... to take care of any other uncaught exception

PRESS ON WITH PIPELINE ... data cleaning is on track ... should work for paper

Do this for all of 1. August

```{r}
fn <- here::here("data","validation-EGLL-APDF-20190801.csv.gz")
egll_arrs_apdf <- read_csv(fn, show_col_types = FALSE)

# safe loop
safe_assign_threshold_time <- possibly(.f = assign_threshold_time, otherwise = "Error")

study_trjs <- arr_trjs |> 
  filter(FLTID %in% egll_arrs_apdf$FLTID) |> 
  mutate(LUID = UID) |> group_by(LUID) |> 
  group_modify(.f = ~safe_assign_threshold_time(.x,.my_thr = my_thr)) |> 
  ungroup() |> select(-LUID)
```

NEXT-NEXT
* calculate characteristics
  + time-to-go
  + distance-to-go
  + crossings (e.g for ASMA)

* save out mst/enriched trajectory