### 2023-03-25 x_validation-EGLL.qmd

---
title: "Untitled"
format: html
---

Check for landing sequence sample from APDF

```{r}
library(tidyverse)

# test sample 
ptn <- "2019-08-01"

apdf <- read_csv("./data/validation-EGLL-2019.csv.gz", show_col_types = FALSE)

arrs <- apdf |> filter(PHASE == "ARR") |> 
  mutate(across(.cols = contains("TIME"), .fns = lubridate::dmy_hms)) |> 
  filter(grepl(pattern = ptn, x = MVT_TIME)) |> 
  mutate(BIN = lubridate::floor_date(MVT_TIME, unit = "15 min"))
```
Runway usage

```{r}
arrs |> ggplot() +
  geom_point(aes(x = MVT_TIME, y = RWY), alpha = 0.3)
```

```{r}
tmp <- arrs |> 
  group_by(BIN, RWY) |> summarise(N = n())

tmp |> 
  ggplot() + geom_step(aes(x = BIN, y = N))
```

peak sessions 10 per 15 mins

```{r}
peak_bins <- tmp |> filter(N >=10)
peak_arrs <- arrs |> filter(BIN %in% peak_bins$BIN) |> arrange(RWY, MVT_TIME)
peak_arrs
```

peak_wave = peak_arrs$FLTID[1:10]

```{r}
fn <- here::here("data","validation-EGLL-APDF-201908-01.csv.gz")
write_csv(arrs, file = fn)
```