### 2023-03-27 x_bra-airport-geoserver.qmd

---
title: "BRA airport data"
format: html
---

BRA airport geo data

based on coordination

* go to geo server https://geoaisweb.decea.mil.br/geoserver/web/wicket/bookmarkable/org.geoserver.web.demo.MapPreviewPage?0
* search for 
  + airport
  rwydirection
  + runway_v2

select file type to download files.

Stored in data/bra.

```{r}
library(tidyverse)

bra_apt_arp <- readr::read_csv(here::here("data","bra","airport.csv"), show_col_types = FALSE)
bra_rwys    <- readr::read_csv(here::here("data","bra","runway_v2.csv"), show_col_types = FALSE)
bra_thrs    <- readr::read_csv(here::here("data","bra","rwydirection.csv"), show_col_types = FALSE)

bra_study_apts <- bra_apt_arp |> 
  filter(localidade_id %in% c("SBGR", "SBSP","SBRJ")) |> 
  select(ICAO = localidade_id, APT_NAME = nome, LAT = latitude_dec, LON = longitude_dec, ELEV = elevacao, airport_pk) |> mutate(ELEV = coerce_meter_to_feet(ELEV))
```

```{r}
check_bra_rwys <- bra_rwys |> filter(airport_pk %in% bra_study_apts$airport_pk) |> glimpse()
```

```{r}
check_bra_thrs <- check_bra_rwys |> filter(runway_pk %in% check_bra_rwys$runway_pk) |> glimpse()
```