### 2023-03-27 x_bra-check.qmd

---
title: "Brazil"
format: html
---

```{r}
library(tidyverse)
library(patchwork)
```

Unfortunately https://aisweb.decea.mil.br/ was down (20.03.).    
Google search offered an AIP pdf from 2017 (hope that no major changes took place - check with DECEA).

Get airport information from AIP

```{r}
knitr::include_graphics("./figures/aip-snapshot-SBGR.png")
```

use measurements package to calculate decimal degrees (rounded to 6 digits)
measurements::conv_unit("046 28 23", "deg_min_sec", "dec_deg")

SBGR: ARP 232608S / 0462823W

bboxfinder - box lng/lat - min/max
bounding box SBGR: -46.500975,-23.460242,-46.435372,-23.416276

```{r}
rwys_sbgr <- tribble(
   ~ICAO  , ~REF , ~LAT      , ~LON      , ~NAME    , ~ELEV
  , "SBGR", "ARP", -23.435555, -46.473056, "ARP"    , 2461
  , "SBGR", "09L", -23.434281, -46.483361, "09L/27R", 2444
  , "SBGR", "27R", -23.424894, -46.448619, "09L/27R", 2440
  , "SBGR", "09R", -23.438797, -46.487027, "09R/27L", 2450
  , "SBGR", "27L", -23.431189, -46.458855, "09R/27L", 2445
)

# write_csv(rwys_sbgr, "./data/rwys-SBGR.csv")
```

```{r}
rwys_sbgr  <- read_csv("./data/rwys-SBGR.csv", show_col_types = FALSE)
sbgr_arp   <- rwys_sbgr |> filter(ICAO == "SBGR", REF == "ARP") |> select(LAT, LON)
sbgr_200NM <- trrrj::polygon_at_distance(c(sbgr_arp$LON, sbgr_arp$LAT), d = 205) |> 
  sf::st_as_sf() |> 
  rename(geometry = x) |> 
  mutate(NAME = "SBGR-200")

# cast sf
arp_pt    <- sbgr_arp  |> cast_latlon_to_pts(.drop_coord = FALSE) 
rwys_pts  <- rwys_sbgr |> filter(REF != "ARP") |> cast_latlon_to_pts(.drop_coord = FALSE)
rwys_ls   <- rwys_sbgr |> cast_rwy_ls()

sbgr_centerline_ls <- cast_rwy_ctr_line_ls(rwys_sbgr |> filter(REF != "ARP"))
sbgr_rwy_box       <- airport_threshold_box(rwys_sbgr |> filter(REF != "ARP"))
sbgr_box           <- airport_centerline_box(rwys_sbgr |> filter(REF != "ARP"), .ctrl_length = 15000)
```

```{r}
sbgr_base_plot <- ggplot() +
  geom_sf(data = sbgr_200NM) +
  geom_sf(data = sbgr_box) +
  geom_point(data = sbgr_arp, aes(x = LON, y = LAT))
sbgr_base_plot
```

## Load OSN data - Monday dump

```{}
# load helper scripts to download and store OSN data
source("./R/osn_dumps.R")
source("./R/osn_read_tardump.R")
source("./R/osn_make_nice_names.R")
```
```{}
#| message: false
pth_to_dumps <- "../__DATA/OSN-MondayDumps"
ptn <- "2022-05-09"

# list downloaded dumps
fns <- list.files(path = pth_to_dumps, pattern = ptn, full.names = TRUE)

# load list of chosen dumps
bra  <- fns |> 
  purrr::map_dfr(.f = read_osn_tar_file) |> 
  make_nice_names_osn()
```

## spatial subsetting for SBGR

```{}
spatial_subset <- function(.payload, .outer, ...){
  # check if payload is already a sf object
  # if not, coerce to sf
  if(!is(.payload,"sf")){
    .payload <- .payload |> cast_latlon_to_pts(.drop_coord = FALSE)
  }
  # spatial subsetting on global scale is calc intensive for sf
  # check for bounding box of .outer
  bbox <- sf::st_bbox(.outer)

  # simple pre-filter based on LAT/LON ~ bounding box
  trimmed_data <- .payload |> 
    dplyr::filter(LAT >= bbox$ymin, LAT <= bbox$ymax) |> 
    dplyr::filter(LON >= bbox$xmin, LON <= bbox$xmax)

  # spatial subsetting
  trimmed_data <- trimmed_data[.outer,]
}

trjs_pts <- spatial_subset(bra |> filter(!is.na(LAT)), sbgr_200NM)

# trjs_pts |> sf::st_drop_geometry() |> 
#   arrow::write_parquet("./data/bra-sbgr-20220509.parquet")
```

```{r}
trjs_pts <- arrow::read_parquet("./data/bra-sbgr-20220509.parquet") |> 
  coerce_unixepoch_to_datetime() |> 
  cast_latlon_to_pts(.drop_coord = FALSE)

sbgr_base_plot + geom_sf(data = trjs_pts)
```

Check for coverage close to SBGR

```{r}
arp_sbgr <- rwys_sbgr |> filter(REF == "ARP") |> select(LAT, LON)
apt_tfc  <- trjs_pts |> 
  trrrj::filter_positions_at_range(
    c(arp_sbgr$LON, arp_sbgr$LAT), 0, 15, lon = LON, lat = LAT
    ) |> 
  filter(ALT_B < 4000)

ggplot() + 
  geom_sf(   data = sbgr_rwys_ls, col = "blue" ) +
  geom_sf(   data = sbgr_box, fill = NA) +
  geom_point(data = apt_tfc, aes(x = LON, y = LAT), size = 0.5)
```

Which ids do we see at SBGR?

```{r}
sbgr_uids <- apt_tfc[sbgr_box,] |> distinct(ICAO24, FLTID)
viz_check_vertical(apt_tfc |> filter(ICAO24 %in% sbgr_uids$ICAO24))
```

```{r}
sbgr_trjs <- trjs_pts |> 
  filter(ICAO24 %in% sbgr_uids$ICAO24) |> 
  identify_trajectory_legs() |> 
  fix_invalid_fltid() |> 
  make_uid("2022-05-09")

ggplot() + geom_sf(data = sbgr_trjs)
```

```{r}
picks <- trjs_pts |> filter(FLTID %in% c("TAM3201", "GLO1612","TAM3172", "DLH506"))
viz_check_DAO(picks |> filter(ALT_B < 8000), sbgr_rwys_ls |> sf::st_buffer(dist = 10000)) + geom_sf(data = sbgr_rwys_ls)

viz_check_vertical(picks)
```

```{r}
picks <- trjs_pts |> filter(FLTID %in% c("TAM3201", "GLO1612","TAM3172", "DLH506"))

sbgr_base_plot + geom_sf(data = picks, aes(color = ALT_B))
```

```{r}
trjs_pts |> filter(FLTID == "DLH506") |> ggplot() + geom_point(aes(x = TIME, y = ALT_B))
```

## Subsetting for KATL

```{r}
katl_arp   <- c(-84.4278640,33.6366996)
katl_200NM <- trrrj::polygon_at_distance(katl_arp, d = 205) |> sf::st_as_sf() |> 
  rename(geometry = x) |> 
  mutate(NAME = "KATL-205")

katl_pts <- spatial_subset(bra |> filter(!is.na(LAT)), katl_200NM)
katl_pts |> sf::st_drop_geometry() |> arrow::write_parquet("./data/usa-katl-20220509.parquet")
```

```{}
# just a black ball = fully covered
ggplot() +
  geom_sf(data = katl_200NM) +
  geom_sf(data = katl_pts)
```

```{r}
check_flight <- katl_pts |> filter(ICAO24 %in% c("a3b4b5"))

p1 <- check_flight |> 
  ggplot() +
  geom_sf(data = katl_200NM) +
  geom_point(aes(x = LON, y = LAT, color = ALT_B))

p2 <- check_flight |> 
  ggplot() +
  geom_point(aes(x = TIME, y = ALT_B, color = ALT_B))

p1 + p2 + plot_layout(guides = 'collect')
```