### 2023-03-30 x_KATL.qmd
building study data loop

---
title: "Spacing Deviation at US airports"
format: html
---

```{r}
library(tidyverse)

files_to_sources = list.files(path = "./R", full.names = TRUE)
purrr::walk(files_to_sources, .f = source)
```

working with ATLANTA

```{r}
rwys_katl <- read_csv("./data/rwys-KATL.csv", show_col_types = FALSE)
arp_katl  <- tibble(LAT = 33.6366996,LON = -84.4278640, REF = "ARP", "ELEV" = 1026)

katl_ctrl_ls <- cast_rwy_ctr_line_ls(rwys_katl)
katl_rwy_box <- airport_threshold_box(rwys_katl)
katl_apt_box <- airport_centerline_box(rwys_katl)

katl_200NM <- trrrj::polygon_at_distance(c(arp_katl$LON, arp_katl$LAT), d = 200) |> 
  sf::st_as_sf() |> mutate(REF = "KATL-200NM")

fn <- here::here("data", "usa-katl-20220509.parquet")
katl_trjs <- arrow::read_parquet(fn)

# |> 
#   coerce_unixepoch_to_datetime() |> 
#   identify_trajectory_legs() |> 
#   fix_invalid_fltid() |> 
#   make_uid("2022-05-09")
```

```{r}
ggplot() +
  geom_sf(data = katl_apt_box) +
  geom_sf(data = katl_rwy_box, fill = "lightblue") +
  geom_sf(data = katl_ctrl_ls) +
  geom_point(data = katl_trjs |> filter(ALT_B <= arp_katl$ELEV + 500), aes(x = LON, y = LAT), size = 0.2, alpha = 0.2) +
  coord_sf(xlim = c(-84.8, -84.1), ylim = c(33.5,33.7) ) +
  labs(x = NULL, y = NULL, subtitle = "KATL - Atlanta Hartsfield") +
  theme(plot.title.position = "plot")
#ggsave("./figures/KATL-traffic-single-day.png", dpi = 300, width = 8, height = 5)
```

identify traffic arriving at Hartsfiled

```{r}
icaos_seen <- katl_trjs |> filter(ALT_B <= arp_katl$ELEV + 1500) |> icao24_in_polygon(katl_apt_box)

study_trjs <- katl_trjs |> filter(ICAO24 %in% icaos_seen$ICAO24) |> 
  coerce_unixepoch_to_datetime() |> 
  identify_trajectory_legs() |> 
 # fix_invalid_fltid() |> 
  mutate(TID = paste(ICAO24, LEG, sep = "-"))

icaos_seen2 <- study_trjs |> 
  filter(ALT_B <= arp_katl$ELEV + 1500) |> 
  icao24_in_polygon(katl_apt_box, .id = TID) |> 
  heuristic_DAO()

heuristic_arrs <- icaos_seen2 |> filter(DAO == "A")

study_trjs <- study_trjs |> filter(TID %in% heuristic_arrs$TID)

arrivals <- study_trjs |> 
  assign_landing_runway(
      .arp = arp_katl
    , .rwy_centerline_ls = katl_ctrl_ls 
    , .upper_bound = 10
    , .group_var = TID
    ) |> 
  filter(N > 5)    # for 10 sec update crossers have typically low number of hits
```
```{r}
itr <- arrivals[1,] |> select(UID = TID, UID_REF = REF)

make_study_arrival_trajectories <- function(.itr, .trjs, .rwys){
  message("\n", itr$UID)
  app_trj <- .trjs |> dplyr::filter(UID == itr$UID)
  app_thr <- .rwys |> dplyr::filter(REF == itr$UID_REF)

  # estimate threshold (aka landing time)
  thr_row <- estimate_threshold_time(app_trj, app_thr) |> 
    select(UID, ICAO24, FLTID, TIME, LAT, LON ) |> 
    mutate(MST = "THR", MST_LABEL = itr$UID_REF)

  # append threshold row and trim trajectory
  app_trj <- app_trj |> bind_rows(thr_row) |> 
    select(UID, ICAO24, FLTID, TIME, LAT, LON, ALT_B, MST, MST_LABEL) |> 
    arrange(TIME) |> 
    rowid_to_column("SEQ_ID")

  # add "distance-to-go" and "time-to-fly" 
  app_trj <- app_trj |> add_flight_distance()
  return(app_trj)
}

bingo <- itr |> make_study_arrival_trajectories(study_trjs |> rename(UID = TID), rwys_katl)
```

run for many, function re-written to add RWY to trjs

```{r}
study_payload <- study_trjs |> 
  select(TID, ICAO24, FLTID, TIME, LAT, LON, ALT_B) |> 
  left_join(arrivals |> select(TID, RWY = REF), by = join_by(TID)) |> 
  rename(UID = TID)

rqq <- study_payload |> 
  mutate(LUID = UID) |> 
  group_by(LUID) |> 
  group_map(.f = ~ safer_make_study_arrival_trajectories2( ., .rwys = rwys_katl))

# check for errors
names(rqq) <- seq_along(rqq)
keep(rqq, ~ ! is.data.frame(.x))
```

```{r}
next_payload <- keep(rqq, ~is.data.frame(.x)) |> bind_rows()
```