### 2023-03-30 x_replicate.qmd

---
title: "Replicate earlier study"
format: html
---

check what we have done in the past

```{r}
library(tidyverse)

# read and load study data
source("~/RProjects/paper-2023-ICNS/R/osn_read_tardump.R")
source("~/RProjects/paper-2023-ICNS/R/osn_make_nice_names.R")
source("~/RProjects/paper-2023-ICNS/R/coerce_meter_to_feet.R")
source("~/RProjects/paper-2023-ICNS/R/coerce_unixepoch_to_datetime.R")
# subset study data ~ jormungand
source("~/RProjects/paper-2023-ICNS/R/spatial_subset_jormungand.R")
source("~/RProjects/paper-2023-ICNS/R/cast_sf_utils.R")

# loop and trajectory processing
source("~/RProjects/paper-2023-ICNS/R/trj_extract_4d_nest.R")
source("~/RProjects/paper-2023-ICNS/R/trj_icao24_in_polygon.R")
source("~/RProjects/paper-2023-ICNS/R/trj_confirm_arrival_at.R")

# runway 
source("~/RProjects/paper-2023-ICNS/R/cast_sf_rwy.R")
```

Load one day of OSN dump

```{}
#| message: false
pth_to_dumps <- "../__DATA/OSN-dumps"
ptn <- "2019-08-01"

# list downloaded dumps
fns <- list.files(path = pth_to_dumps, pattern = ptn, full.names = TRUE)

# load list of chosen dumps
ds  <- fns |> 
  purrr::map_dfr(.f = read_osn_tar_file) |> 
  make_nice_names_osn()
ds <- ds |> 
  mutate(across(.cols = c("ALT_B","ALT_G"), .fns = coerce_meter_to_feet)) 
ds <- ds |> coerce_unixepoch_to_datetime()
```

Subset for study airports
EGLL, EHAM, LSZH

```{}
aip <- read_csv("./data/aip-arp-rwys.csv", show_col_types = FALSE)

apt <- "LSZH"
arp <- filter(aip, ICAO == apt, REF == "ARP") |> rename(ELEV = GEOM_ALT)
jormungand <- trrrj::polygon_at_distance(c(arp$LON, arp$LAT), d = 205)

svs <- spatial_subset_jormungand(ds, jormungand)

arrow::write_parquet(svs |> sf::st_drop_geometry(), paste0("./data/", apt, "-200NM-20190801.feather"))
```

## Load Heathrow

```{r}
apt <- "EGLL"
fn  <- list.files(path = "./data", pattern = paste0(apt,".*\\.feather"), full.names = TRUE)
```

```{r}
study_loop <- function(.apt, .fn, .debug, ...){

  #----------- read airport data -------------------------
  aip <- readr::read_csv("./data/aip-arp-rwys.csv", show_col_types = FALSE) |> 
    dplyr::filter(ICAO == .apt)
  if("GEOM_ALT" %in% names(aip)){
    aip <- aip |> dplyr::rename(ELEV = GEOM_ALT)
    }

  arp     <- aip |> dplyr::filter(ICAO == .apt,   REF == "ARP")
  rwys    <- aip |> dplyr::filter(ICAO == .apt, ! REF == "ARP")
  apt_box <- airport_centerline_box(rwys, ...)

  # ---------------------- load trajectory data --------------------------------
  warning(paste(.apt, " - reading - ", .fn))
  ds <- arrow::read_feather(.fn) %>% 
    tibble::as_tibble() |> 
    #---- enforce above threshold ~ ELEV arp
    dplyr::filter(ALT_B >= arp$ELEV + 300)

  #---------------------- check for local traffic ----------------------------
  # subset icao24s inside airport box
  seen_at_airport <- ds |> 
    dplyr::filter(ALT_B <= arp$ELEV + 4000) |> 
    spatial_subset_jormungand(apt_box, ...) |> 
    dplyr::distinct(ICAO24)

  ds <- ds |> 
    dplyr::filter(ICAO24 %in% seen_at_airport$ICAO24) |> 
    identify_trajectory_legs(.max_gap = 5) |> 
    mutate(UID = paste(ICAO24, LEG, sep = "-"))

  phase <- ds |> extract_4d_nest() |> confirm_arrival_at(apt_box)
}
```

```{r}
gotcha <- study_loop(apt, fn)
```

rq |> 
  # subset for ICAO24s seen at the study airport
  filter(ICAO24 %in% seen_at_airport$ICAO24) |> 
  spatial_subset_jormungand(jormungand) |> 
  # human readable TIME
  coerce_unixepoch_to_datetime() |>
  # clean ALT and trim for airborne part with heuristic fix
  Alt_Outlier_filter(.alt_var = ALT_B) |> 
  fix_altitude_heuristic() |> 
  # assign leg for each airborne segment
  identify_trajectory_legs() |> 
  # force single FLTID here to construct niche UIDs
  group_by(ICAO24, LEG) |> fix_invalid_fltid() |> ungroup() |> 
  # construct UID
  mutate(UID = paste(ICAO24, LEG, sep = "-"))