### 2023-04-01 x_replicate.qmd
getting data prep in place

---
title: "Replicate earlier study"
format: html
---

check what we have done in the past

```{r}
library(tidyverse)

# read and load study data
source("~/RProjects/paper-2023-ICNS/R/osn_read_tardump.R")
source("~/RProjects/paper-2023-ICNS/R/osn_make_nice_names.R")
source("~/RProjects/paper-2023-ICNS/R/coerce_meter_to_feet.R")
source("~/RProjects/paper-2023-ICNS/R/coerce_unixepoch_to_datetime.R")
# subset study data ~ jormungand
source("~/RProjects/paper-2023-ICNS/R/spatial_subset_jormungand.R")
source("~/RProjects/paper-2023-ICNS/R/cast_sf_utils.R")

# loop and trajectory processing
source("~/RProjects/paper-2023-ICNS/R/trj_extract_4d_nest.R")
source("~/RProjects/paper-2023-ICNS/R/trj_icao24_in_polygon.R")
source("~/RProjects/paper-2023-ICNS/R/trj_confirm_arrival_at.R")
source("~/RProjects/paper-2023-ICNS/R/trj_check_orphans.R")

# trajectory cleaning
source("~/RProjects/paper-2023-ICNS/R/trj_fix_altitude.R")
source("~/RProjects/paper-2023-ICNS/R/altitude_outlier_filter.R")

# runway 
source("~/RProjects/paper-2023-ICNS/R/cast_sf_rwy.R")

source("~/RProjects/paper-2023-ICNS/R/assign_threshold_time2.R")
source("~/RProjects/paper-2023-ICNS/R/identify_threshold_overhead.R")
source("~/RProjects/paper-2023-ICNS/R/add_bearing_to_target.R")
source("~/RProjects/paper-2023-ICNS/R/estimate_threshold_time.R")
source("~/RProjects/paper-2023-ICNS/R/identify_threshold_overhead.R")
source("~/RProjects/paper-2023-ICNS/R/add_landing_time_to_nested_data.R")
```

Load one day of OSN dump

```{}
#| message: false
pth_to_dumps <- "../__DATA/OSN-dumps"
ptn <- "2019-08-01"

# list downloaded dumps
fns <- list.files(path = pth_to_dumps, pattern = ptn, full.names = TRUE)

# load list of chosen dumps
ds  <- fns |> 
  purrr::map_dfr(.f = read_osn_tar_file) |> 
  make_nice_names_osn()
ds <- ds |> 
  mutate(across(.cols = c("ALT_B","ALT_G"), .fns = coerce_meter_to_feet)) 
ds <- ds |> coerce_unixepoch_to_datetime()
```

Subset for study airports
EGLL, EHAM, LSZH

```{}
aip <- read_csv("./data/aip-arp-rwys.csv", show_col_types = FALSE)

apt <- "LSZH"
arp <- filter(aip, ICAO == apt, REF == "ARP") |> rename(ELEV = GEOM_ALT)
jormungand <- trrrj::polygon_at_distance(c(arp$LON, arp$LAT), d = 205)

svs <- spatial_subset_jormungand(ds, jormungand)

arrow::write_parquet(svs |> sf::st_drop_geometry(), paste0("./data/", apt, "-200NM-20190801.feather"))
```

## Load Heathrow

```{r}
apt <- "EGLL"
fn  <- list.files(path = "./data", pattern = paste0(apt,".*\\.feather"), full.names = TRUE)
```

```{r}
study_loop <- function(.apt, .fn, .debug, ...){

  #----------- read airport data -------------------------
  aip <- readr::read_csv("./data/aip-arp-rwys.csv", show_col_types = FALSE) |> 
    dplyr::filter(ICAO == .apt)
  if("GEOM_ALT" %in% names(aip)){
    aip <- aip |> dplyr::rename(ELEV = GEOM_ALT)
    }

  arp     <- aip |> dplyr::filter(ICAO == .apt,   REF == "ARP")
  rwys    <- aip |> dplyr::filter(ICAO == .apt, ! REF == "ARP")
  rwy_centerline_ls <- cast_rwy_ctr_line_ls(rwys)
  apt_box <- airport_centerline_box(rwys, ...)

  # ---------------------- load trajectory data --------------------------------
  warning(paste(.apt, " - reading - ", .fn))
  ds <- arrow::read_feather(.fn) %>%
    tibble::as_tibble() |>
    #---- enforce above threshold ~ ELEV arp
    dplyr::filter(ALT_B >= arp$ELEV + 300) |>
     fix_altitude_heuristic() |>
     Alt_Outlier_filter(.alt_var = ALT_B)

  #---------------------- check for local traffic ----------------------------
  # subset icao24s inside airport box
  seen_at_airport <- ds |> 
    dplyr::filter(ALT_B <= arp$ELEV + 4000) |> 
    spatial_subset_jormungand(apt_box, ...) |> 
    dplyr::distinct(ICAO24)

  ds <- ds |> 
    dplyr::filter(ICAO24 %in% seen_at_airport$ICAO24) |> 
    identify_trajectory_legs(.max_gap = 5) |> 
    mutate(UID = paste(ICAO24, LEG, sep = "-"))

  phase <- ds |> 
    extract_4d_nest() |> 
    confirm_arrival_at(apt_box) |> 
    check_arrival_orphan()

  # # add landing time
  # with_ldg_time <-  phase |> 
  #   dplyr::mutate(
  #     study_trj = 
  #     case_when(
  #       ARR & ! ORPHAN ~ pmap(
  #             .l = list(data, UID, RWY)
  #           , .f = ~ add_landing_time_to_nested_data(..1, ..2, ..3, rwys))) 
  #     , TRUE ~ "Departure"
  #   )
}
```

```{r}
gotcha <- study_loop(apt, fn)
```

```{r}
test_landing_time <- function(phase, rwys = rwys_egll){

  arrs <- phase |> filter(ARR & !ORPHAN)
  rest <- phase |> filter(! UID %in% arrs$UID )

  with_ldg_time <-  arrs |>
    dplyr::mutate(
      study_trj = pmap(
              .l = list(data, UID, RWY)
            , .f = ~ add_landing_time_to_nested_data(..1, ..2, ..3, rwys)
            )
      ) |> 
   # dplyr::mutate(study_trj = map(.x = study_trj, .f = ~ interpolate_lin_4D(.x, .source = "OSN"))) |> 
    dplyr::mutate(ALDT = map(.x = study_trj, .f = ~ filter(.x, MST == "ALDT") |> pull(TIME))) |> tidyr::unnest(ALDT)

  return(bind_rows(with_ldg_time, rest))
}

(
  mickey <- gotcha |> test_landing_time()
)
```

```{r}
study_sequences <- mickey |> filter(ARR, !ORPHAN) |> select(UID, RWY, ALDT) |> 
  arrange(RWY, ALDT, UID) |> 
  group_by(RWY) |> 
  mutate(LEAD_DELTA = difftime(ALDT, lag(ALDT, default = first(ALDT)), units = "sec"))
study_sequences
```