### 2023-04-04 study_pipeline.qmd
slowly getting going

---
title: "Industrialise Pipeline"
format: pdf
---

```{r}
list_my_functions <- list.files(path = "./R", pattern = "\\.R", full.names = TRUE)
list_my_functions |> purrr::walk(.f = ~ source(.x))
```

## Overview

This Quarto document provides a staged approach to generating the data for the ICNS2023 paper.

1. load and extract test data for the studied European airports
2. .... generate mickey & store interim results
3. ... do results

## Load and extract study data

```{r}
#| message: false
load_osn_dumps <- function(.ptn, .pth_to_dumps){
    # list downloaded dumps
    fns <- list.files(path = .pth_to_dumps, pattern = .ptn, full.names = TRUE)

    # load list of chosen dumps
    ds  <- fns |> 
      purrr::map_dfr(.f = read_osn_tar_file) |> 
      make_nice_names_osn()
    ds <- ds |> 
      dplyr::mutate(
            dplyr::across(
                  .cols = c("ALT_B","ALT_G")
                , .fns  = coerce_meter_to_feet)
            ) 
    ds <- ds |> coerce_unixepoch_to_datetime()
    return(ds)
}

extract_and_save_jormungand_data <- function(
    .apt, .range = 205, .ds, .save_path = "./data-study/"
    , .save_label = ptn
    , .aip_file = "./data/aip-arp-rwys.csv", ...
    ){
  # load AIP data for study  
  aip <- read_csv(.aip_file, show_col_types = FALSE)
  # extract airport ARP
  arp <- filter(aip, ICAO == .apt, REF == "ARP") |> dplyr::rename(ELEV = GEOM_ALT)
  jormungand <- establish_radial_jormungand(arp, .range) |> 
    dplyr::mutate(REF = paste(ICAO,REF,.range, sep = "-")) |> 
    dplyr::select(ICAO, REF, geometry)

  # subset study data for range
  svs <- spatial_subset_jormungand(.ds, jormungand)

  # write out study data by airport range
  fn_out <- paste0(.save_path, .apt, "-200NM-", .save_label, ".parquet")
  message(paste0("writing ", fn_out))
  arrow::write_parquet(svs |> sf::st_drop_geometry(), fn_out)
}

extract_and_save_study_dataset_for_airport <- function(
    .study_apts
    ,.ptn, .pth_to_dumps, .range = 205, ...
){
  # get osn dump

}

```

Loaded: 
 2019-08-01, 02, 03, 04, 05, 06, 07
 2022-08-01, ... 07, 08, 09

```{r}
# European study data
pth_to_dumps <- "../__DATA/OSN-dumps"
ptn <- "2022-08-09"
study_apts <- c("EGLL","EDDM","EDDF", "LSZH","EHAM")

 osn <- load_osn_dumps(ptn, pth_to_dumps)
  #subset for study airports
  study_apts |> 
    purrr::walk( .f = ~ extract_and_save_jormungand_data(.x, .ds = osn) )
```